Model-based variable selection for regression with highly correlated variables.

### 2014-08-27 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@820 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@821 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@823 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@828 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\paragraph{Further notations} In the following, we will note also $\boldsymbol{J}_r=(J_{r}^1,\ldots,J_r^{d_r})$ the $d_r$-uple of all the response variable (to be not confused with the corresponding set $J_r$ previously defined), $\boldsymbol{J}_p=(J_{p}^1,\ldots,J_p^{d_r})$ the $d_r$-uple of all the predictors for all the sub-regressions, $\boldsymbol{d}_p=(d_p^1,\ldots,d_p^{d_{r}})$ the associated number of predictors and $\boldsymbol{S}=(\boldsymbol{J}_r,\boldsymbol{J}_p)$ the global {\it model} of all the sub-regressions. As more compact notations, we define also $\boldsymbol{X}_r=\boldsymbol{X}^{J_{r}}$ the whole set of response covariates and also $\boldsymbol{X}_f=\boldsymbol{X}^{J_{f}}$ the {\it all} other covariates, denominating now as {\it free} covariates, including these ones used as predictor covariates in $\boldsymbol{J}_p$ or not. An illustration of all these notations is displayed through an example in Section~\ref{sec:tradeoff}. The parameters are also stacked together: $\boldsymbol{\alpha}=(\boldsymbol{\alpha}_1,\ldots,\boldsymbol{\alpha}_{d_r})$ denotes the global coefficient of sub-regressions and $\boldsymbol{\sigma}^2=(\sigma^2_1,\ldots,\sigma^2_{d_r})$ denotes the corresponding global variance.

@@

\paragraph{Qualitative illustration} Noting $\boldsymbol{\beta}^*=(\boldsymbol{\beta}_r^{*'},\boldsymbol{\beta}_f^{*'})'$ the vector of coefficients of the same dimension as $\boldsymbol{\beta}$ with $\boldsymbol{\beta}_r^*={\bf 0}$, (\ref{eq:modexpl}) can be rewritten
\begin{equation}
	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}=\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{X}_r\boldsymbol{\beta}_r^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl2}
\end{equation}
The {\sc ols} estimate of $\boldsymbol{\beta}^*$ is then given by
	\begin{equation}
		\hat{\boldsymbol{\beta}}^*_f=\left(\boldsymbol{X}_f'\boldsymbol{X}_f \right) ^{-1}\boldsymbol{X}_f'\boldsymbol{Y} \quad \mbox{and} \quad \hat{\boldsymbol{\beta}}^*_r = {\bf 0}.
	\end{equation}
As usual, the {\sc ols} estimate $\hat{\boldsymbol{\beta}}^*$ of $\boldsymbol{\beta}^*$ is unbiased but, however, contrary to $\hat{\boldsymbol{\beta}}$, it could be a biased estimate of $\boldsymbol{\beta}$ since
		\begin{equation}
			\mathbb{E}(\hat{\boldsymbol{\beta}}_f^*)=\boldsymbol{\beta}_f+\sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j \quad \textrm{and}\quad \mathbb{E}(\hat{\boldsymbol{\beta}}_r^*)=\boldsymbol{0}.
		\end{equation}
In return, its variance could be reduced compared to this one of $\hat{\boldsymbol{\beta}}$ given in (\ref{eq:varOLS}) as soon as values of $\sigma_j$ are small enough (it means strong correlations in sub-regressions) as we can see in the following expression
		\begin{equation}
			\operatorname{Var}(\hat{\boldsymbol{\beta}}_f^*)= (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2)(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1} \quad \textrm{and} \quad\operatorname{Var}(\hat{\boldsymbol{\beta}}_r^*)= \boldsymbol{0}. \label{eq:varOLS*}
		\end{equation}
Indeed, no correlations between covariates $\boldsymbol{X}_f$ imply that the matrix $\boldsymbol{X}_f' \boldsymbol{X}_f$ could be sufficiently better conditioned than the matrix $\boldsymbol{X}' \boldsymbol{X}$ involved in (\ref{eq:varOLS}) to balance the added variance $\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2$ in (\ref{eq:varOLS*}). This bias-variance trade off can be resumed by the Mean Squared Error (\textsc{mse}) associated to both estimates:
	\begin{eqnarray}
			\textsc{mse}(\hat{\boldsymbol{\beta}})&=& \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}),
			 \\
			\textsc{mse}(\hat{\boldsymbol{\beta}}^*)&=& \parallel\sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j \in I_2}\sigma^2_{j}\beta_{j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}).
	\end{eqnarray}

@@

We define a {\it global} neighbourhood space $\mathcal{V}(\boldsymbol{S})$ of $\boldsymbol{S}$ composed by the following four {\it specific} neighbourhood spaces $\mathcal{V}(\boldsymbol{S})=\mathcal{V}_{r+}(\boldsymbol{S})\cup\mathcal{V}_{r-}(\boldsymbol{S})\cup\mathcal{V}_{p+}(\boldsymbol{S})\cup\mathcal{V}_{p-}(\boldsymbol{S})$ described below:
	\begin{itemize}
	\item {\bf Adding a sub-regression}: a new sub-regression with only one predictor covariate is added to $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{r+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})^{1,\ldots,d_r}=(\boldsymbol{J}_r,\boldsymbol{J}_p), \tilde{J}_r^{d_r+1}\in J_f, \tilde{J}_p^{d_r+1}=\{j\}, j\in J_f \right\}.
\end{equation}	
		\item {\bf Removing a sub-regression}: a sub-regression  is removed from $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{r-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})=(\boldsymbol{J}_r,\boldsymbol{J}_p)^{\{1,\ldots,d_r\}\backslash j}, j\in\{1,\ldots,d_r\} \right\}.
\end{equation}	
	\item {\bf Adding a predictor covariate}: a predictor covariate is added to one sub-regression of $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{p+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{J}_p^{j}=\tilde{J}_p^{j} \cup \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
\end{equation}
	\item {\bf Removing a predictor covariate}: a predictor covariate is removed from one sub-regression of $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{p-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{J}_p^{j}=\tilde{J}_p^{j} \backslash \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \right\}.
\end{equation}	
	\end{itemize}

>>>>

\paragraph{Further notations} In the following, we will note also $\boldsymbol{J}_r=(J_{r}^1,\ldots,J_r^{d_r})$ the $d_r$-uple of all the response variable (to be not confused with the corresponding set $J_r$ previously defined), $\boldsymbol{J}_p=(J_{p}^1,\ldots,J_p^{d_r})$ the $d_r$-uple of all the predictors for all the sub-regressions, $\boldsymbol{d}_p=(d_p^1,\ldots,d_p^{d_{r}})$ the associated number of predictors and $\boldsymbol{S}=(\boldsymbol{J}_r,\boldsymbol{J}_p)$ the global {\it model} of all the sub-regressions. As more compact notations, we define also $\boldsymbol{X}_r=\boldsymbol{X}^{J_{r}}$ the whole set of response covariates and also $\boldsymbol{X}_f=\boldsymbol{X}^{J_{f}}$ the {\it all} other covariates, denominating now as {\it free} covariates, including those used as predictor covariates in $\boldsymbol{J}_p$. An illustration of all these notations is displayed through an example in Section~\ref{sec:tradeoff}. The parameters are also stacked together: $\boldsymbol{\alpha}=(\boldsymbol{\alpha}_1,\ldots,\boldsymbol{\alpha}_{d_r})$ denotes the global coefficient of sub-regressions and $\boldsymbol{\sigma}^2=(\sigma^2_1,\ldots,\sigma^2_{d_r})$ denotes the corresponding global variance.

@@

\paragraph{Qualitative illustration} Noting $\boldsymbol{\beta}^*=(\boldsymbol{\beta}_r^{*'},\boldsymbol{\beta}_f^{*'})'$ the vector of coefficients of the same dimension as $\boldsymbol{\beta}$ with $\boldsymbol{\beta}_r^*={\bf 0}$, (\ref{eq:modexpl}) can be rewritten
\begin{equation}
	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}=\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{X}_r\boldsymbol{\beta}_r^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl2}
\end{equation}
The {\sc ols} estimate of $\boldsymbol{\beta}^*$ is then given by
	\begin{equation}
		\hat{\boldsymbol{\beta}}^*_f=\left(\boldsymbol{X}_f'\boldsymbol{X}_f \right) ^{-1}\boldsymbol{X}_f'\boldsymbol{Y} \quad \mbox{and} \quad \hat{\boldsymbol{\beta}}^*_r = {\bf 0}.
	\end{equation}
As usual, the {\sc ols} estimate $\hat{\boldsymbol{\beta}}^*$ of $\boldsymbol{\beta}^*$ is unbiased but, however, contrary to $\hat{\boldsymbol{\beta}}$, it could be a biased estimate of $\boldsymbol{\beta}$ since
		\begin{equation}
			\mathbb{E}(\hat{\boldsymbol{\beta}}_f^*)=\boldsymbol{\beta}_f+\sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j \quad \textrm{and}\quad \mathbb{E}(\hat{\boldsymbol{\beta}}_r^*)=\boldsymbol{0}.
		\end{equation}
In return, its variance could be reduced compared to this one of $\hat{\boldsymbol{\beta}}$ given in (\ref{eq:varOLS}) as soon as values of $\sigma_j$ are small enough (it means strong correlations in sub-regressions) as we can see in the following expression
		\begin{equation}
			\operatorname{Var}(\hat{\boldsymbol{\beta}}_f^*)= (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2)(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1} \quad \textrm{and} \quad\operatorname{Var}(\hat{\boldsymbol{\beta}}_r^*)= \boldsymbol{0}. \label{eq:varOLS*}
		\end{equation}
Indeed, no correlations between covariates $\boldsymbol{X}_f$ imply that the matrix $\boldsymbol{X}_f' \boldsymbol{X}_f$ could be sufficiently better conditioned than the matrix $\boldsymbol{X}' \boldsymbol{X}$ involved in (\ref{eq:varOLS}) to balance the added variance $\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2$ in (\ref{eq:varOLS*}). This bias-variance trade off can be resumed by the Mean Squared Error (\textsc{mse}) associated to both estimates:
	\begin{eqnarray}
			\textsc{mse}(\hat{\boldsymbol{\beta}})&=& \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}),
			 \\
			\textsc{mse}(\hat{\boldsymbol{\beta}}^*)&=& \parallel\sum_{j =1}^{d_r}\sigma_{j}\beta_{J_r^j} \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}).
	\end{eqnarray}

@@

We define a {\it global} neighbourhood space $\mathcal{V}(\boldsymbol{S})$ of $\boldsymbol{S}$ composed by the following four {\it specific} neighbourhood spaces $\mathcal{V}(\boldsymbol{S})=\mathcal{V}_{r+}(\boldsymbol{S})\cup\mathcal{V}_{r-}(\boldsymbol{S})\cup\mathcal{V}_{p+}(\boldsymbol{S})\cup\mathcal{V}_{p-}(\boldsymbol{S})$ described below:
	\begin{itemize}
	\item {\bf Adding a sub-regression}: a new sub-regression with only one predictor covariate is added to $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{r+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})^{1,\ldots,d_r}=(\boldsymbol{J}_r,\boldsymbol{J}_p), \tilde{J}_r^{d_r+1}\in J_f, \tilde{J}_p^{d_r+1}=\{j\}, j\in J_f \right\}.
\end{equation}	
		\item {\bf Removing a sub-regression}: a sub-regression  is removed from $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{r-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, (\tilde{\boldsymbol{J}_r},\tilde{\boldsymbol{J}_p})=(\boldsymbol{J}_r,\boldsymbol{J}_p)^{\{1,\ldots,d_r\}\backslash j}, j\in\{1,\ldots,d_r\} \right\}.
\end{equation}	
	\item {\bf Adding a predictor covariate}: a predictor covariate is added to one sub-regression of $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{p+}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r, \tilde{J_p}^{\{1,\dots,d_r \}\setminus \{j\}}=J_p^{\{1,\dots,d_r \}\setminus \{j\}} ,\tilde{J}_p^{j}=J_p^{j} \cup \{h\}, j\in \{1,\ldots,d_r\}, h\in J_f \setminus J_p^j\right\}.
\end{equation}
	\item {\bf Removing a predictor covariate}: a predictor covariate is removed from one sub-regression of $\boldsymbol{S}$
\begin{equation}
\mathcal{V}_{p-}(\boldsymbol{S}) = \left\{\tilde{\boldsymbol{S}}: \tilde{\boldsymbol{S}}\in\mathcal{S}, \tilde{\boldsymbol{J}_r}=\boldsymbol{J}_r,\tilde{J_p}^{\{1,\dots,d_r \}\setminus \{j\}}=J_p^{\{1,\dots,d_r \}\setminus \{j\}} , \tilde{J}_p^{j}=J_p^{j} \backslash \{h\}, j\in \{1,\ldots,d_r\}, h\in  J_p^j \right\}.
\end{equation}	
	\end{itemize}
